---
title: "Stats 140SL Assignment 4 Week 3"
author: "Team Members: Kim Jin, Charles Liu, Zoe Wang, Amarissa Mases, Wendy Kha, Jinah Weon"
date: "1/28/2021"
output: pdf_document
---

# Statement of Problem
The Coronavirus 2019 disease pandemic has caused an immense crisis in both health and the economy. From the spread of the infection, the pandemic has impacted people's lives, pushing hospital systems to their capacities, and has created economic deterioration.  

How has the COVID19 pandemic and (its) politics effected the U.S. economy?



# Formal Statistical Hypothesis
The opinion (belief) that "red" states have conferred vulnerability to death by COVID-19 among their populations.



# Analysis (bivariate is perfectly acceptable)


```{r, warning=FALSE, message=FALSE, echo=FALSE}
## Loading Necessary Packages
library(readr)

## Loading Data
setwd(getwd())
covid_deaths <- read_csv("EconomicTracker-main/data/COVID - State - Daily.csv")
## Edit Data to remove "."
covid_deaths$death_number <- ifelse(covid_deaths$death_count == ".",
                                    0, covid_deaths$death_count)
# Create it as numeric values
covid_deaths$num_death_number <- as.numeric(covid_deaths$death_number)

# Amarissa's Code
red_states <- c(1, 2, 5, 12, 16, 18, 19, 20, 21, 22, 28, 29, 30, 31, 37,38, 39, 40, 45, 46, 47, 48, 49, 54, 56) #fips of all states whose popular vote 
covid_deaths$pol_aff <- "blue" #start off with all states as blue
covid_deaths$pol_aff <- ifelse(covid_deaths$statefips %in% red_states, "red", covid_deaths$pol_aff)
# table(covid_deaths$statefips, covid_deaths$pol_aff) #you can check that states were correctly categorized here
covid_deaths$pol_aff <- factor(covid_deaths$pol_aff)

# Death Rate
covid_deaths$death_perct <- ifelse(covid_deaths$death_rate == ".",
                                    0, covid_deaths$death_rate)
covid_deaths$death_perct <- as.numeric(covid_deaths$death_perct)

# New Death Rate
covid_deaths$new_death_perct <- ifelse(covid_deaths$new_death_rate == ".",
                                    0, covid_deaths$new_death_rate)
covid_deaths$new_death_perct <- as.numeric(covid_deaths$new_death_perct)

# New Death Count
covid_deaths$new_death_num <- ifelse(covid_deaths$new_death_count == ".",
                                    0, covid_deaths$new_death_count)
covid_deaths$new_death_num <- as.numeric(covid_deaths$new_death_num)
```

```{r}
# Death Count w/ StateFIP Codes
fit1 <- lm(covid_deaths$num_death_number ~ covid_deaths$statefips, 
          data = covid_deaths)

### All below are blue and red variables

# Death Count
fit2 <- lm(covid_deaths$num_death_number ~ covid_deaths$pol_aff, data = covid_deaths)

# Death Rate
fit3 <- lm(covid_deaths$death_perct ~ covid_deaths$pol_aff, 
          data = covid_deaths)

# New Death Rate
fit4 <- lm(covid_deaths$new_death_perct ~ covid_deaths$pol_aff, 
          data = covid_deaths)

# New Death Count
fit5 <- lm(covid_deaths$new_death_num ~ covid_deaths$pol_aff, 
          data = covid_deaths)

r1 <- summary(fit1)$r.squared
r2 <- summary(fit2)$r.squared
r3 <- summary(fit3)$r.squared
r4 <- summary(fit4)$r.squared
r5 <- summary(fit5)$r.squared

mtx <- matrix(c(r1,r2,r3,r4,r5))
colnames(mtx) <- "R-Squared Values"
rownames(mtx) <- c("r1","r2","r3","r4","r5")
mtx
```



# Conclusion
"type here"



# Limitations of the Analysis
One of the limitation of analyzing COVID-19 data would be uncertainty in the number of cases (e.g. mortality figure might include only deaths occur in hospitals(not other locations such as home).



# Contributions
"type here"







