---
title: "Stats 101C HW 7"
author: "Charles Liu (304804942)"
date: "12/2/2020"
output: 
     pdf_document:
         latex_engine: xelatex
---

# Loading Necessary Packages
```{r, warning=FALSE, message=FALSE}
library(ISLR)
library(plotrix)
library(e1071)
```


# Problem 1 (Exercise 9.7.2)
We have seen that in p = 2 dimensions, a linear decision boundary takes the form $\beta_{0}+\beta_{1}X_{1}+\beta_{2}X_{2}=0$. We now investigate a non-linear decision boundary.

(a) Sketch the curve $(1+X_{1})^{2}+(2-X_{2})^{2}=4$.
```{r}
plot(NA, NA, xlim = c(-4, 1), ylim = c(-1, 5), 
     main = "Problem 1(a) Plot", xlab = "X1", ylab = "X2")
draw.circle(x = -1, y = 2, radius = 1)
```


(b) On your sketch, indicate the set of points for which $(1+X_{1})^{2}+(2-X_{2})^{2}>4$, as well as the set of points for which $(1+X_{1})^{2}+(2-X_{2})^{2}\le4$.
```{r}
plot(NA, NA, xlim = c(-4, 1), ylim = c(-1, 5), 
     main = "Problem 1(b) Plot", xlab = "X1", ylab = "X2")
draw.circle(x = -1, y = 2, radius = 1)
points(-1, 2, pch = 16)
points(-4, 2, pch = 16, col = "red")
legend(-4, 5, legend = c("less than or equal to 4", "greater than 4"),
       pch = 16, col = c("black", "red"), cex = 0.6)
```


(c) Suppose that a classifier assigns an observation to the blue class if $(1+X_{1})^{2}+(2-X_{2})^{2}>4$, and to the red class otherwise. To what class is the observation (0, 0) classified? (−1, 1)? (2, 2)? (3, 8)? 
```{r}
# (0,0) = 5 -> blue
# (-1,1) = 1 -> red
# (2,2) = 9 -> blue
# (3,8) = 52 -> blue
plot(NA, NA, xlim = c(-4, 4), ylim = c(-2, 9), 
     main = "Problem 1(c) Plot", xlab = "X1", ylab = "X2")
draw.circle(x = -1, y = 2, radius = 1)
points(0, 0, pch = 16, col = "blue")
points(-1, 1, pch = 16, col = "red")
points(2, 2, pch = 16, col = "blue")
points(3, 8, pch = 16, col = "blue")
legend(-4, 8, legend = c("greater than 4", "Otherwise"),
       pch = 16, col = c("blue", "red"), cex = 0.8)
```


(d) Argue that while the decision boundary in (c) is not linear in terms of $X_{1}$ and $X_{2}$, it is linear in terms of $X_{1}$, $X_{1}^{2}$, $X_{2}$, and $X_{2}^{2}$.

**ANSWER:** We can simply do this by expanding on the equation from $(1+X_{1})^{2}+(2-X_{2})^{2}=4$ to $X_{1}^{2}+X_{2}^{2}+2X_{1}-4X_{2}+5=4$ and finally we get $X_{1}^{2}+X_{2}^{2}+2X_{1}-4X_{2}+1=0$ (expanded). This shows us that the expanded equation is linear and has terms $X_{1}$, $X_{1}^{2}$, $X_{2}$, and $X_{2}^{2}$.





# Problem 2 (Exercise 9.7.5)
We have seen that we can fit an SVM with a non-linear kernel in order to perform classification using a non-linear decision boundary.We will now see that we can also obtain a non-linear decision boundary by performing logistic regression using non-linear transformations of the features.

(a) Generate a data set with n = 500 and p = 2, such that the observations belong to two classes with a quadratic decision boundary between them. For instance, you can do this as follows:
```{r}
x1 <- runif (500) -0.5
x2 <- runif (500) -0.5
y <- 1*( x1^2-x2^2 > 0)
```


(b) Plot the observations, colored according to their class labels. Your plot should display $X_{1}$ on the x-axis, and $X_{2}$ on the y-axis.
```{r}
plot(x1[y == 0], x2[y == 0], 
     main = "Problem 2(b) Plot", xlab = "X1", ylab = "X2", 
     col = "blue", pch = 2)
points(x1[y == 1], x2[y == 1], col = "green", pch = 3)
```


(c) Fit a logistic regression model to the data, using $X_{1}$ and $X_{2}$ as predictors.
```{r}
lm_2c <- glm(y ~ x1 + x2, family = "binomial")
summary(lm_2c)
```


(d) Apply this model to the training data in order to obtain a predicted class label for each training observation. Plot the observations, colored according to the predicted class labels. The decision boundary should be linear.
```{r}
train_df <- data.frame(x1 = x1, x2 = x2, y = y)
lm_prob <- predict(lm_2c, train_df, type = "response")
lm_predict <- ifelse(lm_prob > 0.5, 1, 0)
lm_one <- train_df[lm_predict == 1, ]
lm_zero <- train_df[lm_predict == 0, ]
plot(NA, NA, xlim = c(-0.5, 0.5), ylim = c(-0.5, 0.5),
     main = "Problem 2(d) Plot", xlab = "X1", ylab = "X2")
points(lm_zero$x1, lm_zero$x2, col = "blue", pch = 2)
points(lm_one$x1, lm_one$x2, col = "green", pch = 3)
```


(e) Now fit a logistic regression model to the data using non-linear functions of $X_{1}$ and $X_{2}$ as predictors (e.g. $X_{1}^{2}$ , $X_{1}*X_{2}$, $log(X_{2})$, and so forth).
```{r, warning=FALSE}
lm_2e <- glm(y ~ poly(x1, 2) + poly(x2, 2) + I(x1 * x2), 
             data = train_df, family = "binomial")
summary(lm_2e)
```


(f) Apply this model to the training data in order to obtain a predicted class label for each training observation. Plot the observations, colored according to the predicted class labels. The decision boundary should be obviously non-linear. If it is not, then repeat (a)-(e) until you come up with an example in which the predicted class labels are obviously non-linear.
```{r}
train_df <- data.frame(x1 = x1, x2 = x2, y = y)
lm_prob <- predict(lm_2e, train_df, type = "response")
lm_predict <- ifelse(lm_prob > 0.5, 1, 0)
lm_one <- train_df[lm_predict == 1, ]
lm_zero <- train_df[lm_predict == 0, ]
plot(NA, NA, xlim = c(-0.5, 0.5), ylim = c(-0.5, 0.5),
     main = "Problem 2(f) Plot", xlab = "X1", ylab = "X2")
points(lm_zero$x1, lm_zero$x2, col = "blue", pch = 2)
points(lm_one$x1, lm_one$x2, col = "green", pch = 3)
```


(g) Fit a support vector classifier to the data with $X_{1}$ and $X_{2}$ as predictors. Obtain a class prediction for each training observation. Plot the observations, colored according to the predicted class labels.
```{r}
svm_2g <- svm(as.factor(y) ~ x1 + x2, train_df, 
              kernel = "linear", scale = FALSE,
              cost = 10, type = "C")
svm_pred <- predict(svm_2g, train_df)
svm_one <- train_df[svm_pred == 1, ]
svm_zero <- train_df[svm_pred == 0, ]
plot(NA, NA, xlim = c(-0.5, 0.5), ylim = c(-0.5, 0.5),
     main = "Problem 2(g) Plot", xlab = "X1", ylab = "X2")
points(svm_zero$x1, svm_zero$x2, col = "blue", pch = 2)
points(svm_one$x1, svm_one$x2, col = "green", pch = 3)
```


(h) Fit a SVM using a non-linear kernel to the data. Obtain a class prediction for each training observation. Plot the observations, colored according to the predicted class labels.
```{r}
svm_2h <- svm(as.factor(y) ~ x1 + x2, train_df)
svm_pred <- predict(svm_2h, train_df)
svm_one <- train_df[svm_pred == 1, ]
svm_zero <- train_df[svm_pred == 0, ]
plot(NA, NA, xlim = c(-0.5, 0.5), ylim = c(-0.5, 0.5),
     main = "Problem 2(g) Plot", xlab = "X1", ylab = "X2")
points(svm_zero$x1, svm_zero$x2, col = "blue", pch = 2)
points(svm_one$x1, svm_one$x2, col = "green", pch = 3)
```


(i) Comment on your results.

**COMMENTS:** We can see that non-linear kernel SVM's are good at locating non-linear boundaries compared to linear kernel SVM's. The results from our logistic regression method is similar to the results of a non-linear kernel SVM's, but we have to create quadratic terms. Creating the quadratic terms could prove to be difficult and non-efficient. Thus, we are better off using non-linear kernel SVM's for this case.





# Problem 3 (Exercise 9.7.8)
This problem involves the OJ data set which is part of the ISLR package.

(a) Create a training set containing a random sample of 800 observations, and a test set containing the remaining observations.
```{r}
data(OJ)
set.seed(1)
train <- sample(nrow(OJ), 800)
oj_train <- OJ[train, ]
oj_test <- OJ[-train, ]
```


(b) Fit a support vector classifier to the training data using cost=0.01, with Purchase as the response and the other variables as predictors. Use the summary() function to produce summary statistics, and describe the results obtained.
```{r}
oj_svm <- svm(Purchase ~ ., data = oj_train, 
              kernel = "linear", scale = FALSE, cost = 10)
summary(oj_svm)
```


(c) What are the training and test error rates?
```{r}
# Create the tables and predictions
oj_train_pred <- predict(oj_svm, oj_train)
oj_train_table <- table(oj_train$Purchase, oj_train_pred)
oj_test_pred <- predict(oj_svm, oj_test)
oj_test_table <- table(oj_test$Purchase, oj_test_pred)

# Show tables
oj_train_table
oj_test_table

# Calculate the error rates for training and test
oj_train_error <- (65+71) / (420+65+71+244)
oj_test_error <- (14+29) / (154+14+29+73)

# Compare the error rates
compare_matrix <- matrix(c(oj_train_error, oj_test_error))
colnames(compare_matrix) <- "Error Rates"
rownames(compare_matrix) <- c("Training", "Test")
compare_matrix
# 17% error rate for training
# 16% error rate for test
```


(d) Use the tune() function to select an optimal cost. Consider values in the range 0.01 to 10.
```{r}
optimal_cost <- tune(svm, Purchase ~ ., data = oj_train, 
                     kernel = "linear", scale = FALSE,
                     ranges = list(cost = 10^seq(-2, 1, by = 0.5)))
# Optimal Cost
optimal_cost$best.parameters$cost
```

**ANSWER:** We find the best parameter for optimal cost is 0.3162278.


(e) Compute the training and test error rates using this new value for cost.
```{r}
# Create the tables and predictions
oj_svm_best <- svm(Purchase ~ ., kernel = "linear", data = oj_train, 
                       cost = optimal_cost$best.parameters$cost)
oj_train_pred_best <- predict(oj_svm_best, oj_train)
oj_train_table_best <- table(oj_train$Purchase, oj_train_pred_best)
oj_test_pred_best <- predict(oj_svm_best, oj_test)
oj_test_table_best <- table(oj_test$Purchase, oj_test_pred_best)

# Show tables
oj_train_table_best
oj_test_table_best

# Calculate the error rates for training and test
oj_train_error_best <- (62+71) / (423+62+71+244)
oj_test_error_best <- (13+29) / (155+13+29+73)

# Compare the error rates
compare_matrix <- matrix(c(oj_train_error_best, oj_test_error_best))
colnames(compare_matrix) <- "Error Rates"
rownames(compare_matrix) <- c("Training", "Test")
compare_matrix
# 16.625% error rate for training
# 15.556% error rate for test
```


(f) Repeat parts (b) through (e) using a support vector machine with a radial kernel. Use the default value for gamma.
```{r}
### part (b)
set.seed(1)
oj_svm_radial <- svm(Purchase ~ ., data = oj_train, kernel = "radial")
summary(oj_svm_radial)


### part (c)
# Create the tables and predictions
oj_train_pred_radial <- predict(oj_svm_radial, oj_train)
oj_train_table_radial <- table(oj_train$Purchase, oj_train_pred_radial)
oj_test_pred_radial <- predict(oj_svm_radial, oj_test)
oj_test_table_radial <- table(oj_test$Purchase, oj_test_pred_radial)

# Show tables
oj_train_table_radial
oj_test_table_radial

# Calculate the error rates for training and test
oj_train_error_radial <- (44+77) / (441+44+77+238)
oj_test_error_radial <- (33+17) / (151+17+33+69)

# Compare the error rates
compare_matrix <- matrix(c(oj_train_error_radial, oj_test_error_radial))
colnames(compare_matrix) <- "Error Rates"
rownames(compare_matrix) <- c("Training", "Test")
compare_matrix
# 15.125% error rate for training
# 18.519% error rate for test


### part (d)
optimal_cost_radial <- tune(svm, Purchase ~ ., data = oj_train, 
                     kernel = "radial", scale = FALSE,
                     ranges = list(cost = 10^seq(-2, 1, by = 0.5)))
# Optimal Cost
optimal_cost_radial$best.parameters$cost


### part (e)
# Create the tables and predictions
oj_svm_best_radial <- svm(Purchase ~ ., kernel = "radial", data = oj_train, 
                       cost = optimal_cost_radial$best.parameters$cost)
oj_train_pred_best_radial <- predict(oj_svm_best_radial, oj_train)
oj_train_table_best_radial <- table(oj_train$Purchase,
                                    oj_train_pred_best_radial)
oj_test_pred_best_radial <- predict(oj_svm_best_radial, oj_test)
oj_test_table_best_radial <- table(oj_test$Purchase, oj_test_pred_best_radial)

# Show tables
oj_train_table_best_radial
oj_test_table_best_radial

# Calculate the error rates for training and test
oj_train_error_best_radial <- (43+73) / (442+43+73+242)
oj_test_error_best_radial <- (14+36) / (154+14+36+66)

# Compare the error rates
compare_matrix <- matrix(c(oj_train_error_best_radial,
                           oj_test_error_best_radial))
colnames(compare_matrix) <- "Error Rates"
rownames(compare_matrix) <- c("Training", "Test")
compare_matrix
# 14.5% error rate for training
# 18.519% error rate for test
```


(g) Repeat parts (b) through (e) using a support vector machine with a polynomial kernel. Set degree=2. 
```{r}
### part (b)
set.seed(1)
oj_svm_poly <- svm(Purchase ~ ., data = oj_train, 
                     kernel = "polynomial", degree = 2)
summary(oj_svm_poly)


### part (c)
# Create the tables and predictions
oj_train_pred_poly <- predict(oj_svm_poly, oj_train)
oj_train_table_poly <- table(oj_train$Purchase, oj_train_pred_poly)
oj_test_pred_poly <- predict(oj_svm_poly, oj_test)
oj_test_table_poly <- table(oj_test$Purchase, oj_test_pred_poly)

# Show tables
oj_train_table_poly
oj_test_table_poly

# Calculate the error rates for training and test
oj_train_error_poly <- (110+36) / (110+36+205+449)
oj_test_error_poly <- (45+15) / (45+15+153+57)

# Compare the error rates
compare_matrix <- matrix(c(oj_train_error_poly, oj_test_error_poly))
colnames(compare_matrix) <- "Error Rates"
rownames(compare_matrix) <- c("Training", "Test")
compare_matrix
# 18.25% error rate for training
# 22.22% error rate for test


### part (d)
#optimal_cost_poly <- tune(svm, Purchase ~ ., data = oj_train, 
#                     kernel = "polynomial", degree = 2, scale = FALSE,
#                     ranges = list(cost = 10^seq(-2, 1, by = 0.5)))
# Optimal Cost
#optimal_cost_poly$best.parameters$cost
# Difficult to load but it came out as cost = 10 (best)

### part (e)
# Create the tables and predictions
oj_svm_best_poly <- svm(Purchase ~ ., kernel = "polynomial", degree = 2,
                        data = oj_train,
                        cost = 10)
oj_train_pred_best_poly <- predict(oj_svm_best_poly, oj_train)
oj_train_table_best_poly <- table(oj_train$Purchase,
                                    oj_train_pred_best_poly)
oj_test_pred_best_poly <- predict(oj_svm_best_poly, oj_test)
oj_test_table_best_poly <- table(oj_test$Purchase, oj_test_pred_best_poly)

# Show tables
oj_train_table_best_poly
oj_test_table_best_poly

# Calculate the error rates for training and test
oj_train_error_best_poly <- (38+82) / (447+38+82+233)
oj_test_error_best_poly <- (14+37) / (154+14+37+65)

# Compare the error rates
compare_matrix <- matrix(c(oj_train_error_best_poly,
                           oj_test_error_best_poly))
colnames(compare_matrix) <- "Error Rates"
rownames(compare_matrix) <- c("Training", "Test")
compare_matrix
# 15% error rate for training
# 18.889% error rate for test
```


(h) Overall, which approach seems to give the best results on this data?

**ANSWER:** We can see that radial kernel with the optimal cost of 10 yielded the lowest error rates with $training = 14.5\%$ and $test = 18.519\%$. For this data, we should use the Radial Kernel with its optimal cost.