---
title: "Stats_101B_HW_5_Charles_Liu (Dis. 3A)"
author: "Charles Liu (304804942)"
date: "5/25/2020"
output: pdf_document
---

# Loading Necessary Packages:
```{r, warning=FALSE, message=FALSE}
library(DoE.base)
library(unrepx)
```



# Problem 1: (Exercise 7.6)
```{r}
# Set up the problem
response1 <- c(7,9,34,55,16,20,40,60,
               8,10,32,50,18,21,44,61,
               8,12,35,52,15,22,45,65,
               6,10,30,53,15,20,41,63)

k=5
contrast1 <- contr.FrF2(2^(k))
colnames(contrast1) <- c("A", "B","AB", "C", "AC","BC","ABC","D","AD",
"BD","ABD","CD", "ACD","BCD","ABCD","E","AE","BE","ABE","CE","ACE","BCE",
"ABCE","DE","ADE","BDE","ABDE","CDE","ACDE","BCDE","ABCDE")
contrast1

# Estimated Effects w/out Blocks
effect1 <- t(response1)%*%contrast1/(16*1)
names(effect1) <- c("A", "B","AB", "C", "AC","BC","ABC","D","AD",
"BD","ABD","CD", "ACD","BCD","ABCD","E","AE","BE","ABE","CE","ACE","BCE",
"ABCE","DE","ADE","BDE","ABDE","CDE","ACDE","BCDE","ABCDE")
effect1

# Adding in Blocks
block1 <- contrast1[,c("ABCDE")]
da1 <- data.frame(contrast1,block1,response1)
model1 <- aov(response1~factor(A)*factor(B)*factor(C)*factor(D)*factor(E)+factor(block1),data=da1)
summary(model1)

# Estimated effects w/ Blocks
effect1_a <- t(response1)%*%cbind(contrast1)/(16*1)
colnames(effect1_a)[colnames(effect1_a) == "ABCDE"] <- "ABCDE+block1"
effect1_a # -0.1875 is the estimate of (block effect + effect of interaction ABCDE)
effect1_b <- 2*coef(lm(response1~A*B*C*D*E+block1,data=da1))[-1]
effect1_b

# Estimates of effects for treatment factors.
# In other words, remove the highest-order interaction and the blocking factor, as they are confounded.
effect1_c <- effect1_a[setdiff(names(effect1_a),c("block","A:B:C:D:E"))]
effect1_c
# using Lenth's method
t.tests1 <- eff.test(effect1, method = "Lenth")
t.tests1[,-5]


# Using normal probability plot
qqnorm(effect1)
text(qqnorm(effect1)$x,qqnorm(effect1)$y, names(effect1))
qqline(effect1)
```

ANOVA does not print the ABCDE interaction, since it is confounded with blocking. Also, the estimated block effect includes ABCDE interaction effect. (-0.1875) is the estimate of (block effect + effect of interaction ABCDE). We see that "A, B, C, AB" are significant factors from our T-test using Lenth's Method. We create our Normal Q-Q Plot and see that "A, B, C, AB" are significant.



# Problem 2: (Exercise 7.24)
```{r}
mx1_b <- matrix(names(effect1_b), nrow=8, ncol=4, byrow=FALSE)
mx1_b
```

The answer is (d) "BE" is in the same block as "ACDE". We can see that from the 6th row that "B:E" is along the same row as "A:C:D:E".



# Problem 3:
```{r}
# Choice 1: B_1 = {1,2,3}, B_2 = {4,5,6,7}, B_3 = {3,4,5}
# Choice 2: B_1 = {1,2,3,4}, B_2 = {3,4}, B_3 = {5,6,7}
# Choice 3: B_1 = {1,2,5,6}, B_2 = {1,2,3,4}, B_3 = {1,3,5,7}

# C1 --> (B_1)(B_2)(B_3) --> (1,2,3)(4,5,6,7)(3,4,5) = (1,2,6,7) 
# C2 --> (B_1)(B_2)(B_3) --> (1,2,3,4)(3,4)(5,6,7) = (1,2,5,6,7) 
# C3 --> (B_1)(B_2)(B_3) --> (1,2,5,6)(1,2,3,4)(1,3,5,7) = (4,6,7) 

# 3 Block Effects!
# In Conclusion,...

# C3 < C1 < C2
```

Since we are using 3 Block Effects, we can see that Choice 3 (C3) is the worst one of the choices. After that, we see that Choice 1 (C1) is the next worst with a confounding effect of (1,2,6,7). Lastly, we see that Choice 2 (C2) has the best confounding effect of (1,2,5,6,7) out of all the choices given.