---
title: "stats_101b_hw3_Charles_Liu"
author: "Charles Liu (304804942)"
date: "4/23/2020"
output: pdf_document
---

# Loading Necessary Packages:
```{r, message = FALSE}
library(Hmisc)
library(DescTools)
library(tidyverse)
library(crossdes)
```


# Problem 1 (Exercise 4.8)
```{r}
values <- c(13, 22, 18, 39,
          16, 24, 17, 44,
          5, 4, 1, 22)
days <- rep(c("1 day", "2 day", "3 day", "4 day"), 3)
solutions <- c(rep("1", 4), rep("2", 4), rep("3", 4))
milk_data <- data.frame(values = values, solutions = solutions, days = days)
milk_data

m1 <- lm(values ~ days + solutions, data = milk_data)
summary(m1)
anova(m1)
TukeyHSD(aov(m1))
par(mfrow = c(2, 2))
plot(m1)
```

We can see that since the p-values for both Number of Days and Types of Solutions are small (or their F_0 values are very big), we would REJECT the Null Hypothesis. This means that both Days and Solutions have a significant effect on retarding bacteria growth.

We can see that the following pairs are significant from our Tukey Method for Days: 4 day-1 day, 4 day-2 day, & 4 day-3 day. The following pairs are significant from our Tukey Method for Solutions: 3-1 & 3-2. (all our Tukey Method uses a significance level of $\alpha$ = 0.05)

From our plots, we can see it relatively satisfies the Normality asummption, but it does NOT satisfy the Linearity and Homoscedasticity assumptions (Residuals vs. Fitted & Scale-Location plots).



# Problem 2 (Exercise 4.24)
## Problem 2 (done by "hand")
```{r}
letter <- c("A", "B", "D", "C", "E",
            "C", "E", "A", "D", "B",
            "B", "A", "C", "E", "D",
            "D", "C", "E", "B", "A",
            "E", "D", "B", "A", "C")
value <- c(8, 7, 1, 7, 3,
            11, 2, 7, 3, 8,
            4, 9, 10, 1, 5,
            6, 8, 6, 6, 10,
            4, 2, 3, 8, 8)
day <- rep(c("Day 1", "Day 2", "Day 3", "Day 4", "Day 5"), 5)
batch <- c(rep("1", 5), rep("2", 5), rep("3", 5), rep("4", 5), rep("5", 5))
chem_data <- data.frame(letter = letter, value = value, batch = batch, day = day)
chem_data

# I'll be calculating everything by "hand" using Rstudio
total_batch <- chem_data %>% select(batch, value) %>% group_by(batch) %>% summarise(row_total = sum(value))
total_batch

total_day <- chem_data %>% select(day, value) %>% group_by(day) %>% summarise(column_total = sum(value))
total_day

total_letter <- chem_data %>% select(letter, value) %>% group_by(letter) %>% summarise(letter_sum = sum(value))
total_letter

total <- chem_data %>% select(letter, value) %>% summarise(total_values = sum(value)) # 147

# Setting up our paramters:
N <- 25 # (5x5) matrix
n <- 5 # can work for both Day and Batch since it's a (5x5) matrix
p <- 5
CF <- (147^2)/(N)

# Sum of Squares:
row_per_batch <- c(26, 31, 29, 36, 25)
SSR_batch <- ((1/n) * (sum(row_per_batch[1:5]^2))) - CF # SSR_batch

column_per_day <- c(33, 28, 27, 25, 34)
SSR_day <- ((1/n) * (sum(column_per_day[1:5]^2))) - CF # SSR_day

letter_per <- c(42, 28, 44, 17, 16)
SSR_letter <- ((1/n) * (sum(letter_per[1:5]^2))) - CF # SSR_letter

# Created a for() loop for Total Sum of Squares
for(i in 1:25) {
  new_value <- (value[1:25])^2
  SS_all <- sum(new_value)
  return(SS_all)
}

SST <- SS_all - CF # Total Sum of Squares

SSE <- SST - SSR_batch - SSR_day - SSR_letter # Sum of Square of Errors

# Mean Sum of Squares:
MS_batch <- SSR_batch/(p - 1)
MS_day <- SSR_day/(p - 1)
MS_letter <- SSR_letter/(p - 1)
MSE <- SSE/((p - 2) * (p - 1))

# F-Test:
F_batch <- (MS_batch/MSE)
F_day <- (MS_day/MSE)
F_letter <- (MS_letter/MSE) 



# All of our Values for the ANOVA table:
## Degrees of Freedom Values:
(p - 1) # Rows, Columns, Letters
(p - 2)*(p - 1) # SSE
(p^2 - 1) # SST

## Sum of Square Values:
SSR_batch
SSR_day
SSR_letter
SSE
SST

## Mean Sum of Square Values:
MS_batch
MS_day
MS_letter
MSE

## F-test Values:
F_batch
F_day
F_letter
```


## Problem 2 (done by RStudio)
```{r}
# Simple Way by Using lm():
m2 <- lm(value ~ day + batch + letter, data = chem_data)
summary(m2)
anova(m2) # We can see all our answers done by hand are the same!
TukeyHSD(aov(m2))
par(mfrow = c(2, 2))
plot(m2)
```

We can see that only Ingredients (Letters) have a small p-value, so we would REJECT the Null Hypothesis for Ingredients. Therefore, we can say that Ingredients have a significant effect on the Chemical Process. As for Day and Batch, they are both NOT significant, so we would NOT REJECT the Null Hypothesis. For these two, we can say they do NOT play a significant role on the Chemical Process.

There is no significance for Day and Batch under Tukey Method (with a significance level $\alpha$ = 0.05). We can see under Ingredients (Letters), there is significance between the following pairs: D-A, E-A, D-C, & E-C.

From our plots, we can see it relatively satisfies the Normality and Homoscedasticity asummptions, but it does NOT satisfy the Linearity assumptions (Residuals vs. Fitted).



# Problem 3 (Exercise 4.37)
```{r}
standing_times <- c("A", "B", "C", "D", "E",
                    "B", "C", "D", "E", "A",
                    "C", "D", "E", "A", "B",
                    "D", "E", "A", "B", "C",
                    "E", "A", "B", "C", "D")
catalysts <- c("alpha", "beta", "gamma", "delta", "epsilon",
               "gamma", "delta", "epsilon", "alpha", "beta",
               "epsilon", "alpha", "beta", "gamma", "delta",
               "beta", "gamma", "delta", "epsilon", "alpha",
               "delta", "epsilon", "alpha", "beta", "gamma")
amount <- c(26, 16, 19, 16, 13,
           18, 21, 18, 11, 21,
           20, 12, 16, 25, 13,
           15, 15, 22, 14, 17,
           10, 24, 17, 17, 14)
acid_concentration <- rep(c("Acid 1", "Acid 2", "Acid 3", "Acid 4", "Acid 5"), 5)
batches <- c(rep("1", 5), rep("2", 5), rep("3", 5), rep("4", 5), rep("5", 5))
chem_process_data <- data.frame(standing_times = standing_times, catalysts = catalysts, amount = amount, acid_concentration = acid_concentration, batches = batches)
chem_process_data

# Using RStudio to find conclusions (lm(...) method)
m3 <- lm(amount ~ standing_times + catalysts + acid_concentration + batches, data = chem_process_data)
summary(m3)
anova(m3)
TukeyHSD(aov(m3))
par(mfrow = c(2, 2))
plot(m3)
```

We can see that only Standing Times (Latin Letters) have a small p-value, so we would REJECT the Null Hypothesis for Standing Times Therefore, we can say that Standing Times have a significant effect on the Chemical Process. As for Catalysts (Greek Letters), Acid Concentration, and Batches, they are all NOT significant, so we would NOT REJECT the Null Hypothesis. For these three, we can say they do NOT play a significant role on the Chemical Process.

There is no significance for Catalysts (Greek Letters), Acid Concentration, and Batches under Tukey Method (with a significance level $\alpha$ = 0.05). We can see under Standing Times (Latin Letters), there is significance between the following pairs: B-A, D-A, E-A, & E-C.

From our plots, we can see it relatively satisfies the Linearity, Normality, and Homoscedasticity asummptions. Even if lines are not perfectly horizontal, they are mostly horizontal. Therefore, all our assumptions are satisfied.



# Problem 4 (Exercise 4.44)
```{r}
amounts <- c(114, 0, 0, 0, 120, 0, 117,
           126, 120, 0, 0, 0, 119, 0,
           0, 137, 117, 0, 0, 0, 134,
           141, 0, 129, 149, 0, 0, 0,
           0, 145, 0, 150, 143, 0, 0,
           0, 0, 120, 0, 118, 123, 0,
           0, 0, 0, 136, 0, 130, 127)
time <- rep(c("Day 1", "Day 2", "Day 3", "Day 4", "Day 5", "Day 6", "Day 7"), 7)
hardwood_concentration <- c(rep("2%", 7), rep("4%", 7), rep("6%", 7), rep("8%", 7), rep("10%", 7), rep("12%", 7), rep("14%", 7))
paper_data <- data.frame(time = time, hardwood_concentration = hardwood_concentration, amounts = amounts)
paper_data

# Using RStudio to find conclusions (lm(...) method)
m4 <- lm(amounts ~ time + hardwood_concentration, data = paper_data)
summary(m4)
anova(m4)
TukeyHSD(aov(m4))
par(mfrow = c(2, 2))
plot(m4)
```

We can see that none of the factors has a p-value less than 0.05, so they are all NOT significant. We would NOT REJECT the Null Hypothesis, and we can say they do NOT play a significant role on the Strength of Paper.

There is no significance for Time and Hardwood Concentration under Tukey Method (with a significance level $\alpha$ = 0.05).

From our plots, we can see that NONE of our assumptions are satisfied (Linearity, Normality, and Homoscedasticity asummptions).



# Problem 5
```{r}
design_1 <- matrix(c(1, 2, 3, 4, 5,
                     2, 3, 4, 5, 1,
                     3, 4, 5, 1, 3,
                     4, 5, 1, 2, 4), nrow = 5, ncol = 4)

design_2 <- matrix(c(1, 5, 2, 3, 1,
                     2, 1, 3, 4, 2,
                     3, 2, 4, 5, 4,
                     4, 3, 5, 1, 5), nrow = 5, ncol = 4)
```


## 5a)
```{r}
# (a > k); a*gamma = b * k; (a - 1) * lambda = (k - 1) * gamma

design_1
# a = 5, b = 5, k = 4, gamma = 3, lambda = 4
# (i) a * gamma = b * k --> (5) * (3) = (5) * (4) --> NOT TRUE (=/=)
# (ii) (a - 1) * (lambda) = (k - 1) * (gamma) --> (5 - 1) * (4) = (4 - 1) * (3) --> NOT TRUE (=/=)


design_2
# a = 5, b = 5, k = 4, gamma = 4, lambda = 3
# (i) a * gamma = b * k --> (5) * (4) = (5) * (4) --> TRUE (=)
# (ii) (a - 1) * (lambda) = (k - 1) * (gamma) --> (5 - 1) * 3 = (4 - 1) * (4) --> TRUE (=)






# Creating our Incidence Matrices:
incidence_1 <- matrix(c(1, 1, 1, 1, 0,
                        0, 1, 1, 1, 1,
                        1, 0, 1, 1, 1,
                        1, 1, 0, 1, 1,
                        1, 0, 1, 1, 1), nrow = 5, ncol = 5)
incidence_1

incidence_2 <- matrix(c(1, 1, 1, 1, 0,
                        1, 1, 1, 0, 1,
                        0, 1, 1, 1, 1,
                        1, 0, 1, 1, 1,
                        1, 1, 0, 1, 1), nrow = 5, ncol = 5)
incidence_2
```


## 5b)
```{r}
# Testing our Incidence Matrices:
incidence_1 %*% t(incidence_1) # This one is NOT BIBD.
incidence_2 %*% t(incidence_2) # This one is BIBD! We would choose Design 2 for BIBD.


# Double Checking our Work:
isGYD(design_1)
# NOT BIBD

isGYD(design_2)
# Is BIBD
```
