---
title: "Stats 101B HW 2 Lecture 3A"
author: "Charles Liu (304804942)"
date: "4/18/2020"
output: pdf_document
---

# Loading Packages:
```{r}
require(Hmisc)
require(DescTools)
```



# Problem 2:
## Creating Data for Problem 2:
```{r}
observations <- c(24, 28, 37, 30,
                  37, 44, 31, 35,
                  42, 47, 52, 38)
dosages <- c(rep("20g", 4), rep("30g", 4), rep("40g", 4))
drug.data <- data.frame(dosages = dosages, observations = observations)
drug.data
```


## 2b)
```{r}
summary(observations ~ dosages, data = drug.data)

m1 <- lm(observations ~ dosages, data = drug.data)
summary(m1)
aov(m1)
anova(m1)
```


## 2e)
```{r}
MSE_2 <- 32.028
qtukey(1 - 0.05, 3, 12 - 3)/(sqrt(2))
tstat <-(mean(observations[dosages == "30g"]) - 
           mean(observations[dosages == "20g"]))/sqrt(2*MSE_2/4)
tstat # H_0: M_1 = M_2
1-ptukey(tstat*sqrt(2),3, 12-3) # p-value
TukeyHSD(x= aov(m1), 'dosages', conf.level = 0.95)
```

We fail to REJECT the NUll Hypothesis and we can say they are NOT significantly different for the means.


## 2f)
```{r}
res_2 <- m1$residual
fitted_2 <- m1$fitted.values
par(mfrow = c(1, 2))
# normality check
qqnorm(res_2); qqline(res_2)
# no pattern / constant variance check
plot(fitted_2, res_2)
```

All assumptions are satisfied for all the plots.


# Problem 3:
## Creating Data for Problem 3:
```{r}
conductivity <- c(143, 141, 150, 146,
                  152, 149, 137, 143,
                  134, 136, 132, 127,
                  129, 127, 132, 129)
coating_type <- c(rep("1", 4), rep("2", 4), rep("3", 4), rep("4", 4))
tv.data <- data.frame(coating_type = coating_type, conductivity = conductivity)
tv.data
```


## 3b)
```{r}
summary(conductivity ~ coating_type, data = tv.data)
m2 <- lm(conductivity ~ coating_type, data = tv.data)
summary(m2)
aov(m2)
anova(m2)
```


## 3d)
```{r}
MSE_3 <- 19.6875 # from the ANOVA table
mean(conductivity[coating_type == "1"]) - 
  mean(conductivity[coating_type == "4"]) + 
  qt(1 - 0.005, 16 - 4)*sqrt(2*MSE_3/4)

mean(conductivity[coating_type == "1"]) - 
  mean(conductivity[coating_type == "4"]) - 
  qt(1 - 0.005, 16 - 4)*sqrt(2*MSE_3/4)
```


## 3e)
```{r}
TukeyHSD(x= aov(m2), 'coating_type', conf.level = 0.95)
```

We fail to REJECT the NUll Hypothesis and we can say they are NOT significantly different for the means.


## 3f)
```{r}
res_3 <- m2$residual
fitted_3 <- m2$fitted.values
par(mfrow = c(1, 2))
# normality check
qqnorm(res_3); qqline(res_3)
# no pattern / constant variance check
plot(fitted_3, res_3)
```

All assumptions are satisfied for all the plots.