---
title: "Stats 102A - Homework 2 - Output File"
author: "Charles Liu (304804942)"
output: pdf_document
---

Homework questions and prompts copyright Miles Chen, Do not post, share, or distribute without permission.

To receive full credit the functions you write must pass all tests. We may conduct further tests that are not included on this page as well.

# Academic Integrity Statement

By including this statement, I, Charles Liu, declare that all of the work in this assignment is my own original work. At no time did I look at the code of other students nor did I search for code solutions online. I understand that plagiarism on any single part of this assignment will result in a 0 for the entire assignment and that I will be referred to the dean of students.

I did discuss ideas related to the homework with Leah Skelton for Part 4.At no point did I show another student my code, nor did I look at another student's code.


\pagebreak

# Part 1: Board representation

Create a single object called `board` where you store the features of the game board in R.

```{r, error = FALSE}
Board <- list("nrow" = 10, "ncol" = 10, "Chutes" =  '98, 95, 93, 87, 64, 62, 56, 49, 47, 16',
              "Ladders" = '1, 4, 9, 21, 28, 36, 51, 71, 80', "Start Point" = '1', 
              "End Point" = '100')
Board
```

\pagebreak

# Part 2: Plot of Game board

```{r, error = FALSE}
show_board <- function(miniboard = FALSE) {
  if(miniboard == FALSE) {
plot.new()
# Vertical Line
plot.window(xlim = c(0, 55), ylim = c(0, 55))
segments(x0 = seq(from = 5, to = 55, by = 5), x1 = seq(from = 5, to = 55, by = 5), 
         y0 = 0, y1 = 100)
# Horizontal Line
segments(x0 = 0, x1 = 100, 
         y0 = seq(from = 5, to = 55, by = 5), y1 = seq(from = 5, to = 55, by = 5))
text(x = seq(from = 7, to = 55, by = 5), y = 8, label = '1':'10')
text(x = seq(from = 7, to = 55, by = 5), y = 13, label = '20':'11')
text(x = seq(from = 7, to = 55, by = 5), y = 18, label = '21':'30')
text(x = seq(from = 7, to = 55, by = 5), y = 23, label = '40':'31')
text(x = seq(from = 7, to = 55, by = 5), y = 28, label = '41':'50')
text(x = seq(from = 7, to = 55, by = 5), y = 33, label = '60':'51')
text(x = seq(from = 7, to = 55, by = 5), y = 38, label = '61':'70')
text(x = seq(from = 7, to = 55, by = 5), y = 43, label = '80':'71')
text(x = seq(from = 7, to = 55, by = 5), y = 48, label = '81':'90')
text(x = seq(from = 7, to = 55, by = 5), y = 53, label = '100':'91')

arrows(x0 = 5, x1 = 19, 
       y0 = 4, y1 = 25, col = 'green', lwd = 2)
arrows(x0 = 20, x1 = 38, 
       y0 = 5, y1 = 14, col = 'green', lwd = 2)
arrows(x0 = 48, x1 = 52, 
       y0 = 5, y1 = 25, col = 'green', lwd = 2)
arrows(x0 = 5, x1 = 14, 
       y0 = 15, y1 = 30, col = 'green', lwd = 2)
arrows(x0 = 45, x1 = 24, 
       y0 = 15, y1 = 48, col = 'green', lwd = 2)
arrows(x0 = 30, x1 = 23, 
       y0 = 20, y1 = 28, col = 'green', lwd = 2)
arrows(x0 = 55, x1 = 37, 
       y0 = 30, y1 = 37, col = 'green', lwd = 2)
arrows(x0 = 53, x1 = 53, 
       y0 = 40, y1 = 55, col = 'green', lwd = 2)
arrows(x0 = 7, x1 = 7, 
       y0 = 42, y1 = 53, col = 'green', lwd = 2)

arrows(x0 = 18, x1 = 18, 
       y0 = 53, y1 = 40, col = 'red', lwd = 2)
arrows(x0 = 33, x1 = 33, 
       y0 = 53, y1 = 40, col = 'red', lwd = 2)
arrows(x0 = 43, x1 = 43, 
       y0 = 53, y1 = 40, col = 'red', lwd = 2)
arrows(x0 = 23, x1 = 5, 
       y0 = 38, y1 = 33, col = 'red', lwd = 2)
arrows(x0 = 13, x1 = 13, 
       y0 = 38, y1 = 12, col = 'red', lwd = 2)
arrows(x0 = 25, x1 = 43, 
       y0 = 33, y1 = 33, col = 'red', lwd = 2)
arrows(x0 = 48, x1 = 54, 
       y0 = 30, y1 = 10, col = 'red', lwd = 2)
arrows(x0 = 38, x1 = 33, 
       y0 = 30, y1 = 17, col = 'red', lwd = 2)
arrows(x0 = 28, x1 = 33, 
       y0 = 14, y1 = 8, col = 'red', lwd = 2)
}
  else {
    plot.new()
    plot.window(xlim = c(0, 50), ylim = c(0, 50))
    segments(x0 = seq(from = 5, to = 35, by = 5), x1 = seq(from = 5, to = 35, by = 5), 
             y0 = 0, y1 = 50)
    segments(x0 = 0, x1 = 50, 
             y0 = seq(from = 5, to = 40, by = 5), y1 = seq(from = 5, to = 40, by = 5))
    text(x = seq(from = 7, to = 35, by = 5), y = 8, label = '1':'6')
    text(x = seq(from = 7, to = 35, by = 5), y = 13, label = '12':'7')
    text(x = seq(from = 7, to = 35, by = 5), y = 18, label = '13':'18')
    text(x = seq(from = 7, to = 35, by = 5), y = 23, label = '24':'19')
    text(x = seq(from = 7, to = 35, by = 5), y = 28, label = '25':'30')
    text(x = seq(from = 7, to = 35, by = 5), y = 33, label = '36':'31')
    text(x = seq(from = 7, to = 35, by = 5), y = 38, label = '37':'42')
    
    arrows(x0 = 5, x1 = 13, 
           y0 = 5, y1 = 23, col = 'green', lwd = 2)
    arrows(x0 = 23, x1 = 18, 
           y0 = 10, y1 = 27, col = 'green', lwd = 2)
    arrows(x0 = 27, x1 = 33, 
           y0 = 22, y1 = 32, col = 'green', lwd = 2)
    arrows(x0 = 22, x1 = 27, 
           y0 = 30, y1 = 38, col = 'green', lwd = 2)
    
    arrows(x0 = 22, x1 = 32, 
           y0 = 18, y1 = 7, col = 'red', lwd = 2)
    arrows(x0 = 33, x1 = 33, 
           y0 = 30, y1 = 17, col = 'red', lwd = 2)
    arrows(x0 = 12, x1 = 18, 
           y0 = 33, y1 = 33, col = 'red', lwd = 2)
    arrows(x0 = 23, x1 = 12, 
           y0 = 38, y1 = 28, col = 'red', lwd = 2)
}
}

## TEST!!!!!!!!
show_board()
```

\pagebreak

# Part 3: Miniboard

Create the `miniboard` object and create a plot

```{r, error = FALSE}
show_board(miniboard = TRUE)
```

\pagebreak

# Part 4: Verbose output of one single player game

```{r, error = FALSE}
## For Ladders
func1 <- function(newposition) {
  Ladders <- rbind(c(1, 38), c(4, 14), c(9, 31), c(21, 42), c(28, 84), c(36, 44), 
                   c(51, 67), c(71, 91), c(80, 100))
  if(any(Ladders[, 1]) == newposition) {
    row1 <- Ladders[Ladders[,1] == 1]
    newval1 <- row1[2]
  }
  else {
    newval1 <- newposition
  }
  newval1
}

## For Chutes
func2 <- function(newposition) {
  Chutes <- rbind(c(98, 78), c(95, 75), c(93, 73),c(64, 60), c(62, 19), c(56, 53), 
                  c(49, 11), c(47, 26), c(16, 6))
  if(any(Chutes[, 1]) == newposition) {
    row2 <- Chutes[Chutes[,1] == 1]
    newval2 <- row2[2]
  }
  else {
    newval2 <- newposition
  }
  newval2
}

## GAME!
play_solo <- function(x, verbose) {
  # Using by rows for our Chutes or Ladders
  # c("land on", "moved to")
  # Possibly separate Chutes and Ladders
  Ladders <- rbind(c(1, 38), c(4, 14), c(9, 31), c(21, 42), c(28, 84), c(36, 44), 
                   c(51, 67), c(71, 91), c(80, 100))
  Chutes <- rbind(c(98, 78), c(95, 75), c(93, 73),c(64, 60), c(62, 19), c(56, 53), 
                  c(49, 11), c(47, 26), c(16, 6))
  
  newposition <- 0
  oldposition <- 0
  possiblemove <- 0
  turns <- 0
  moves <- 0
  ladders_tally <- 0
  chutes_tally <- 0
  
  if(verbose == TRUE) {
  
    while(all(oldposition < 100)) {
    cat("Turns:", turns + 1, "\n")
    cat("Started at:", oldposition, "\n")
    spin <- sample(1:6, 1)
    cat("Spin Landed On:", spin, "\n")
    possiblemove <- (oldposition + spin)
    if(possiblemove > 100) {
      while(possiblemove > 100) {
        cat("The number is over 100! \n", "Spin Again Please! \n")
        spin <- sample(1:6, 1)
        possiblemove <- oldposition + spin
      }
    }
    newposition <- possiblemove
    temp1 <- newposition
    newposition <- func1(newposition)
    if(temp1 != newposition && temp1 < newposition) {
      ladders_tally <- ladders_tally + 1
      cat("LADDERS!", "\n")
    }
    newposition <- func2(newposition)
    if(temp1 != newposition && temp1 > newposition) {
      chutes_tally <- chutes_tally + 1
      cat("CHUTES!", "\n")
    }
    oldposition <- newposition
    oldposition
    turns <- turns + 1
    cat("Ended at:", oldposition, "\n")
    moves <- append(moves, oldposition)
    cat("\n")
  }
  oldposition
  cat("YOU WIN after", turns, "Turns \n")
  cat("MOVES DONE:", moves, "\n")
  cat("Number of Ladders Done:", ladders_tally, "\n")
  cat("Number of Chutes Done:", chutes_tally, "\n")
  }
  else {
    while(all(oldposition < 100)) {
      spin <- sample(1:6, 1)
      possiblemove <- (oldposition + spin)
      if(possiblemove > 100) {
        while(possiblemove > 100) {
          spin <- sample(1:6, 1)
          possiblemove <- oldposition + spin
        }
      }
      newposition <- possiblemove
      temp1 <- newposition
      newposition <- func1(newposition)
      if(temp1 != newposition && temp1 < newposition) {
        ladders_tally <- ladders_tally + 1
      }
      newposition <- func2(newposition)
      if(temp1 != newposition && temp1 > newposition) {
        chutes_tally <- chutes_tally + 1
      }
      oldposition <- newposition
      oldposition
      turns <- turns + 1
      moves <- append(moves, oldposition)
      cat("\n")
    }
    oldposition
  cat("YOU WIN after", turns, "Turns \n")
  cat("MOVES DONE:", moves, "\n")
  cat("Number of Ladders Done:", ladders_tally, "\n")
  cat("Number of Chutes Done:", chutes_tally, "\n")
  }
}



## TEST!!!!!!!!!!!
play_solo(player1, verbose = FALSE)

play_solo(player1, verbose = TRUE)
```

\pagebreak

# Part 5: Monte Carlo Simulation Study

```{r, error = FALSE, warning = FALSE}
# If I don't use Error = FALSE, then there will be a lot of extra information. I've also used warning = FALSE.
replicate(10000, play_solo(player1, verbose = FALSE))
```


- Create a histogram (breaks = 50) of the turns.

```{r}
func1 <- function(newposition) {
  Ladders <- rbind(c(1, 38), c(4, 14), c(9, 31), c(21, 42), c(28, 84), c(36, 44), 
                   c(51, 67), c(71, 91), c(80, 100))
  if(any(Ladders[, 1]) == newposition) {
    row1 <- Ladders[Ladders[,1] == 1]
    newval1 <- row1[2]
  }
  else {
    newval1 <- newposition
  }
  newval1
}

func2 <- function(newposition) {
  Chutes <- rbind(c(98, 78), c(95, 75), c(93, 73),c(64, 60), c(62, 19), c(56, 53), 
                  c(49, 11), c(47, 26), c(16, 6))
  if(any(Chutes[, 1]) == newposition) {
    row2 <- Chutes[Chutes[,1] == 1]
    newval2 <- row2[2]
  }
  else {
    newval2 <- newposition
  }
  newval2
}

Ladders <- rbind(c(1, 38), c(4, 14), c(9, 31), c(21, 42), c(28, 84), c(36, 44), 
                   c(51, 67), c(71, 91), c(80, 100))
Chutes <- rbind(c(98, 78), c(95, 75), c(93, 73),c(64, 60), c(62, 19), c(56, 53), 
                  c(49, 11), c(47, 26), c(16, 6))

gnewposition <- 0
oldposition <- 0
possiblemove <- 0
turns <- 0
moves <- 0
ladders_tally <- 0
chutes_tally <- 0

while(all(oldposition < 100)) {
  cat("Turns:", turns + 1, "\n")
  cat("Started at:", oldposition, "\n")
  spin <- sample(1:6, 1)
  cat("Spin Landed On:", spin, "\n")
  possiblemove <- (oldposition + spin)
  if(possiblemove > 100) {
    while(possiblemove > 100) {
      cat("The number is over 100! \n", "Spin Again Please! \n")
      spin <- sample(1:6, 1)
      possiblemove <- oldposition + spin
    }
  }
  newposition <- possiblemove
  temp1 <- newposition
  newposition <- func1(newposition)
  if(temp1 != newposition && temp1 < newposition) {
    ladders_tally <- ladders_tally + 1
    cat("LADDERS!", "\n")
  }
  newposition <- func2(newposition)
  if(temp1 != newposition && temp1 > newposition) {
    chutes_tally <- chutes_tally + 1
    cat("CHUTES!", "\n")
  }
  oldposition <- newposition
  oldposition
  turns <- turns + 1
  cat("Ended at:", oldposition, "\n")
  moves <- append(moves, oldposition)
  cat("\n")
}


hist(moves, breaks = 50, rep(moves, 10000))
```


- Find the minimum number of turns. How many times out of 10,000 did a game finish with the minimum number of turns?

```{r}
# I know it'll take about 7 turns for the minimum. I just can't figure out how to code for the minimum number of turns due to my turn count being inside my while loops.
```

- Find the maximum number of turns.
```{r}
max(moves, rep(moves, 10000))
```


- What is the median number of turns?
```{r}
median(moves, rep(moves, 10000))
```

- What is the mean number of turns?
```{r}
mean(moves, trim = 0.5, rep(moves, 10000))
```

- What proportion of games take 100 or more turns to complete?
```{r}
if(length(moves) > 100) {
  moves
}
```

- What proportion of games take 10 or fewer turns to complete?
```{r}
if(length(moves) < 10) {
  moves
}
```

- What proportion of games utilize ladder 9 (the shortcut to win on space 80)?
```{r}
# I'm not quite sure how to do this one.
```

- Create a barplot of the relative frequency of how often each chute is utilized. (Number the chutes in order based on their starting square. The chute with lowest starting number, 16 to 6, is chute 1. The chute going from 98 to 78 is chute 10.)
```{r}
barplot(Chutes, xlab = "Chutes", ylab = "Frequency of Chutes")
```

- Create a barplot of the relative frequency of how often each ladder is utilized. (Number the ladders in order based on their starting square. The ladder with lowest starting number, 1 to 38, is ladder 1. The Ladder going from 80 to 100 is ladder 9.)
```{r}
barplot(Ladders, xlab = "Ladders", ylab = "Frequency of Ladders")
```
